% resume.tex
%
% (c) 2002 Matthew Boedicker <mboedick@mboedick.org> (original author) http://mboedick.org
% (c) 2003 David J. Grant <dgrant@ieee.org> http://www.davidgrant.ca
% (c) 2007 Todd C. Miller <Todd.Miller@courtesan.com> http://www.courtesan.com/todd
%
%This work is licensed under the Creative Commons Attribution-NonCommercial-ShareAlike License. To view a copy of this license, visit http://creativecommons.org/licenses/by-nc-sa/1.0/ or send a letter to Creative Commons, 559 Nathan Abbott Way, Stanford, California 94305, USA.

\documentclass[letterpaper,10pt]{article}
\usepackage[us,12hr]{datetime} % `us' makes \today behave as usual in TeX/
%-----------------------------------------------------------

%\usepackage{timestamp}
%\usepackage[empty]{fullpage}
\usepackage{amsmath}
\usepackage{color}
\usepackage{url}
\definecolor{mygrey}{gray}{0.80}
\raggedbottom
%\raggedright
\setlength{\tabcolsep}{0in}

\title{Bennett Clayton}
% Adjust margins to 0.5in on all sides
%\addtolength{\oddsidemargin}{-1in}
%\addtolength{\evensidemargin}{-4in}
%\addtolength{\textwidth}{1.0in}
%\addtolength{\topmargin}{0in}
%\addtolength{\textheight}{1.0in}

%-----------------------------------------------------------
%Custom commands
\newcommand{\resitem}[1]{\item #1 \vspace{-2pt}}
\newcommand{\resheading}[1]{{\large \colorbox{mygrey}{\begin{minipage}{\textwidth}{\textbf{#1 \vphantom{p\^{E}}}}\end{minipage}}}}
\newcommand{\ressubheading}[4]{
\begin{tabular*}{7.0in}{l@{\extracolsep{\fill}}r}
		\textbf{#1} & #2 \\
		\textit{#3} & \textit{#4} \\
\end{tabular*}\vspace{-6pt}}
%-----------------------------------------------------------

\setlength{\oddsidemargin}{0.3cm} \setlength{\topmargin}{0in}
\setlength{\textwidth}{6.4in}
\setlength{\textheight}{8.5in} \setlength{\parindent}{-0.1in}

\begin{document}
\hfill{\today\ }
\vspace{1cm}
\begin{center}
{\textbf{\LARGE Bennett Clayton}}\\
\vspace{.2cm}
\textbf{Email:} {\color{blue} bgclayto@tamu.edu}
%\Large August 2011
\end{center}


\vspace{0.1in}

\resheading{Education}

\begin{itemize}
\item \textbf{Texas A\&M University}, \hfill Fall 2016 -
	Present \\PhD in Mathematics 
\vspace{0.1cm}

\item \textbf{University of North Carolina at Charlotte},
	\hfill Spring 2014 \\Masters of Science in
	Mathematics 
\vspace{0.1cm}

\item \textbf{University of North Carolina at Charlotte},
	\hfill Fall 2011 \\Bachelor of Science in
	Mathematics
\vspace{0.1cm}
\textit{Cum Laude}

\vspace{0.5cm}
\end{itemize}

\resheading{Publications}
\begin{itemize}
	\item B. Clayton, J.-L. Guermond, M. Maier, B.
		Popov, E. J. Tovar. (Preprint) \textit{Robust
		second-order approximation of the
	compressible Euler equations with an arbitrary
equation of state.} ArXiv (2022).
\url{https://arxiv.org/abs/2207.12832}.  
	\item B. Clayton, J.-L. Guermond, B. Popov.
		\textit{Invariant Domain Preserving
		Approximations for the Euler Equations with
	Tabulated Equation of State.} SIAM Journal on
	Scientific Computing, 44 (2022), pp. A444-A470.
\end{itemize}
\vspace{0.5cm}

\resheading{Research interests}

\begin{itemize}
\item \textbf{Conservation Laws (Euler Equations)}
\item \textbf{Numerical Analysis}
\end{itemize}
\vspace{0.5cm}


\resheading{Current Research}

\begin{itemize}
\item \textbf{Advisors}: Dr. Popov, Dr. Guermond, and Dr.
	Maier. \textbf{Research}: Development of invariant
	domain preserving methods for the Euler equations
	with tabulated EOS. 
\end{itemize}
\vspace{0.5cm}


\resheading{Presentations}
\begin{itemize}
\item \textbf{WCCM-APCOM 2022 (Virtual)} \hfill August
	2022\\ Yokohama, Japan.  Title: \textit{Invariant
		Domain Preserving Methods for the Euler
		equations with a Tabulated Equation of
	State.}
\item \textbf{4th Annual Meeting of the SIAM Texas-Louisiana
		Section} \hfill November 2021\\ South Padre
		Island, TX, USA.  Title:
		\textit{Invariant-Domain Preserving
			Approximation of the Compressible
			Euler Equations with Tabulated
		Equations of State.} 
\item \textbf{Center for Large-Scale Scientific Simulations
		Seminar} \hfill April 2021 \\ Texas A\&M
		University, College Station, TX, USA. Title:
		\textit{Invariant-Domain Preserving
			Approximation of the Compressible
		Euler Equations with Tabulated Equations of
	State.} 
\end{itemize}


\resheading{Teaching Experience}
\vspace{0.4cm}

\large{\textbf{Texas A\&M University}, Instructor of Record
	\hfill Summer 2022

\normalsize

\begin{enumerate}
\item[] \textbf{MATH 140 Mathematics for Business and Social
	Sciences} \hfill Summer 2022 \\ \textit{One class}
\end{enumerate}

\large{\textbf{Texas A\&M University}, Teaching Assistant
\hfill Fall 2017 - Summer 2021} \\

\normalsize

\begin{enumerate}
\item[] \textbf{Numerical Qualifier Preparatory Class} \hfill Summer 2021 \\
\textit{One class} \\
\textit{Duties included: preparing graduate students for their numerical PhD qualifying exam.}
\item[] \textbf{MATH 442 Math Modeling} \hfill Fall 2019 \\
\textit{One class} \\
\textit{Duties included: preparing assignments and projects, grading, teaching Python, \\ and managing a lab.}
\item[] \textbf{Numerical Qualifier Preparatory Class} \hfill Summer 2019 \\
\textit{One class} \\
\textit{Duties included: preparing graduate students for their numerical PhD qualifying exam.}
\item[] \textbf{MATH 417 Numerical Analysis} \hfill Spring 2019 \\
\textit{One section.} \\
\textit{Duties included: teaching Python, recitation, and grading.}
\item[] \textbf{MATH 151 Engineering Calculus I} \hfill Fall 2018 \\
\textit{Three sections.} \\
\textit{Duties included: teaching MATLAB, recitation, quiz preparation, and grading.}
\item[] \textbf{MATH 152 Engineering Calculus II} \hfill Spring 2018 \\
\textit{Three sections.} \\
\textit{Duties included: teaching MATLAB, recitation, and grading.}
\item[] \textbf{MATH 151 Engineering Calculus I} \hfill Fall 2017 \\
\textit{Three sections.} \\
\textit{Duties included: teaching MATLAB, recitation, quiz preparation, and grading.}
\end{enumerate}

\large{\textbf{UNC Charlotte}, Part-time Instructor \hfill Fall 2014 - Spring 2016}
\vspace{0.1cm}

\normalsize

%Courses: 

\begin{itemize}
\item \textbf{MATH 1100 College Algebra} (\textit{Two
	courses}) \hfill Spring 2016 \\ \textbf{MATH 1241
Calculus I} (\textit{One course}) \\ \textbf{MATH 2164
Linear Algebra} (\textit{One course}) 
\end{itemize}

\begin{itemize}
\item \textbf{MATH 1100 College Algebra}  
(\textit{Two courses}) \hfill Fall 2015 \\
\textbf{MATH 1120 Business Calculus}
(\textit{Two courses})
\end{itemize}

\begin{itemize}
\item \textbf{MATH 1120 Business Calculus} 
    (\textit{One course}) \hfill Spring 2015 \\
    \textbf{MATH 1242 Calculus II} (Online)
    (\textit{One course})
\end{itemize}

\begin{itemize}
    \item \textbf{MATH 1120 Business Calculus} (Online)
	  (\textit{One course}) \hfill Fall 2014 \\
    \textbf{MATH 1120 Business Calculus} (\textit{One
    course}) \\
    \textbf{Calculus for Engineering Technology}
    (\textit{Two courses})
\end{itemize}

%\large{\textbf{UNC Charlotte}, Graduate Student \hfill Spring 2014}
%
%\normalsize
%
%\begin{itemize}
%\item \textbf{MATH 1103 Precalculus} \hfill Spring 2014 \\
%(\textit{One course})
%\end{itemize}

\vspace{0.5cm}


\resheading{Outreach}

\begin{itemize}
\item \textbf{Texas Junior Science and Humanities Symposia} \hfill January 2022 \\
(\textit{Judge for Mathematics and Comp. Sci. Group})
\item \textbf{Texas A\&M Integral Bee} \hfill Spring 2019 \\
(\textit{Volunteer})
\item \textbf{Texas A\&M High School Mathematics Contest} \hfill Fall 2016 \\
(\textit{Volunteer Grader})
\end{itemize}
\vspace{0.5cm}


\resheading{Leadership}

\begin{itemize}
    \item \textbf{Society for Industrial and Applied Math} \hfill Fall 2019 - Spring 2020 \\
    (\textit{TAMU Graduate Student Chapter: President})
    \item \textbf{Society for Industrial and Applied Math} \hfill Fall 2018 - Spring 2019 \\
    (\textit{TAMU Graduate Student Chapter: Liason Officer})
\end{itemize}
\vskip 0.5cm


\resheading{Skills}

\begin{itemize}
    \item \textbf{Programming Languages}: Fortran and C++.
    \item \textbf{Other Sofware/Programs}: LaTeX, \texttt{deal.II}, MATLAB
\end{itemize}
\vspace{0.5cm}


\resheading{Software Development} 
\begin{itemize}
\item \textbf{Ryujin}: A high performance multi-threaded
	parallel C++ code based on the \texttt{deal.II}
	finite element library for solving the Euler
	equations. 
\item \textbf{In-house Fortran Code}: A Fortran program for
	solving the Euler equations using continuous finite
	elements. 
\end{itemize}


\resheading{Languages}

\begin{itemize}
\item English: Native language
\item Russian: Intermediate level
\item Korean: Beginner level
\end{itemize}
\vspace{0.5cm}


\resheading{Awards}

\begin{itemize}
\item \textbf{Award for Academic Excellence}   \hfill Spring 2009 \\
UNC Charlotte, Department of Russian Studies 
\end{itemize}
\vspace{0.5cm}

\end{document}
